Dynamics of Laterally Propagating Flames in X-ray Bursts

### 2019-05-14 xrb/paper.tex
fix some equations
add some section assignments

\newcommand{\maestro}{{\sf MAESTRO}}
\newcommand{\castro}{{\sf Castro}}
\newcommand{\pynucastro}{{\sf pynucastro}}

@@

All simulations are performed with the \castro\ hydrodynamics
code~(\citealt{castro}; see also \citealt{astronum:2017} for a recent
description).  We evolve the system of fully compressible Euler
equations for reacting flow:
\begin{eqnarray}
\frac{\partial( \rho X_k)}{\partial t} &=& - \nabla\cdot (\rho \Ub X_k) + \rho \omegadot_k \\
%
\frac{\partial (\rho \Ub)}{\partial t} &=& - \nabla\cdot (\rho \Ub \Ub) - \nabla p +
    \rho \gb \nonumber \\
  &&-2 \rho \Omegab\times \Ub - \Omegab \times (\Omegab \times \rb) \\
%
\frac{\partial (\rho E)}{\partial t} &=& - \nabla \cdot (\rho \Ub E + p\Ub) +
    \nabla \cdot \kth \nabla T + \rho \epsdot +\nonumber\\
  && \rho \Ub \cdot \gb - \rho (\Omegab \cdot \rb)(\Omegab \cdot \Ub) + \rho |\Omegab|^2 (\Ub \cdot \rb)
\end{eqnarray}
Here, $\rho$ is the mass density, $\Ub$ is the velocity, and $E$ is
the specific total energy, which is related to the specific internal
energy as $e = E - |\Ub|^2/2$.  Species are described by mass
fractions, $X_k$ (such that $\sum_k X_k = 1$), and creation rates,
$\omegadot$, and are related to the total specific energy generation
rate, $\epsdot$.  The total mass conservation,
\begin{equation}
\frac{\partial \rho}{\partial t} = - \nabla\cdot (\rho \Ub)
\end{equation}
implies $\sum_k \omegadot_k = 0$.  \castro\ uses an unsplit piecewise
parabolic method (PPM) with characteristic tracing for solving the
hydrodynamics~\cite{ppm,millercolella:2002}, generalized to an
arbitrary equation of state~\cite{zingalekatz}.  Reactions are
incorporated via Strang splitting~\cite{strang:1968}, giving a method that is overall
second-order accurate in space and time.

@@

Since the neutron star rotates, we work in a corotating frame, with
angular velocity $\Omegab$.  Further, for the two-dimensional
simulations presented here, we work in axisymmetric coordinates,
but we advective a third component of velocity, coming out of the
simulation plane, that participates in the Coriolis force (sometimes
described as a 2.5D simulation).  We will take the
\castro\ $x$-coordinate to be the cylindrical radial coordinate with
corresponding velocity $u$, the \castro\ $y$-coordinate to be the
cylindrical vertical coordinate with corresponding velocity $v$, and the
\castro\ $z$-coordinate to be the cylindrical azimuthal coordinate,
with corresponding velocity $w$.  A righthanded coordinate system has
positive $w$ pointing out of the page.  We will take $\Omegab =
\Omega_0 \hat{\bf y}$ for the angular rotation rate, and $\gb = -g
\hat{\bf y}$ for the gravitational acceleration, with $g$ constant.
With these choices, the Coriolis force is:
\begin{equation}
-2\rho \Omegab \times \Ub =
   -2\rho \left ( \Omegab w \hat{x} - \Omegab u \hat{z} \right )
\end{equation}
We will neglect the centrifugal force---with our plane-parallel
geometry, this will act only in the lateral direction, and is not
expected to greatly affect the dynamics.  Carrying the Coriolis force
allows us to capture the geostrophic balance\MarginPar{ref} that sets
up via lateral hydrostatic equilibrium.

Writing the momentum equation in terms of the $u$, $v$, and $w$
components, and neglecting the centrifugal force, we have:
\begin{align}
\frac{\partial (\rho u)}{\partial t} + \nabla \cdot (\rho u \Ub) +
     \frac{\partial p}{\partial x} &= -2\rho \Omega w  \\
%
\frac{\partial (\rho v)}{\partial t} + \nabla \cdot (\rho v \Ub) +
     \frac{\partial p}{\partial y} &= -\rho g \\
%
\frac{\partial (\rho w)}{\partial t} + \nabla \cdot (\rho w \Ub) +
     \cancelto{0}{\frac{\partial p}{\partial z}} &=
    2\rho \Omega u 
\end{align}
where we cancel $\partial p/\partial z$ because there are no
variations in the azimuthal direction.  This allows us to recast the
$w$-velocity equation as a simple advection equation:
\begin{equation}
\frac{\partial w}{\partial t} + \Ub \cdot \nabla w = 2\Omega u
\end{equation}
In our geometry, the flame will propagate from left to right, so $u$
will be positive and the Coriolis force results in $w > 0$ (out of the
simulation plane).  The algorithmic implementation of rotation in
\castro\ is described in \cite{wdmergerI}.

@@

@@

@@

We use aprox13 instead of the small network to understand the
differences.  Energy comes out faster.

\subsection{Sponge}

@@

@@

MHD, radiation transport

@@

\software{AMReX (\url{https://github.com/AMReX-Codes/amrex}),
          Castro \citep{castro},
          GCC (\url{https://gcc.gnu.org/}),
          linux (\url{https://www.kernel.org/}),
          matplotlib (\citealt{Hunter:2007}, \url{http://matplotlib.org/}),
          NumPy \citep{numpy,numpy2},
          python (\url{https://www.python.org/}),
          valgrind \citep{valgrind},
          VODE \citep{vode},
          yt \citep{yt}}

>>>>

\newcommand{\maestro}{{\sf MAESTRO}}
\newcommand{\castro}{{\sf Castro}}
\newcommand{\amrex}{{\sf AMReX}}
\newcommand{\pynucastro}{{\sf pynucastro}}

@@

All simulations are performed with the \castro\ hydrodynamics
code~(\citealt{castro}; see also \citealt{astronum:2017} for a recent
description).  We evolve the system of fully compressible Euler
equations for reacting flow:
\begin{eqnarray}
\frac{\partial( \rho X_k)}{\partial t} &=& - \nabla\cdot (\rho \Ub X_k) + \rho \omegadot_k \\
%
\frac{\partial (\rho \Ub)}{\partial t} &=& - \nabla\cdot (\rho \Ub \Ub) - \nabla p +
    \rho \gb \nonumber \\
  &&-2 \rho \Omegab\times \Ub - \Omegab \times (\Omegab \times \rb) \\
%
\frac{\partial (\rho E)}{\partial t} &=& - \nabla \cdot (\rho \Ub E + p\Ub) +
    \nabla \cdot \kth \nabla T + \rho \epsdot +\nonumber\\
  && \rho \Ub \cdot \gb - \rho (\Omegab \cdot \rb)(\Omegab \cdot \Ub) + \rho |\Omegab|^2 (\Ub \cdot \rb)
\end{eqnarray}
Here, $\rho$ is the mass density, $\Ub$ is the velocity, and $E$ is
the specific total energy, which is related to the specific internal
energy as $e = E - |\Ub|^2/2$.  Species are described by mass
fractions, $X_k$ (such that $\sum_k X_k = 1$), and creation rates,
$\omegadot$, and are related to the total specific energy generation
rate, $\epsdot$.  The total mass conservation,
\begin{equation}
\frac{\partial \rho}{\partial t} = - \nabla\cdot (\rho \Ub)
\end{equation}
implies $\sum_k \omegadot_k = 0$.  \castro\ uses an unsplit piecewise
parabolic method (PPM) with characteristic tracing for solving the
hydrodynamics~\cite{ppm,millercolella:2002}, generalized to an
arbitrary equation of state~\cite{zingalekatz}.  Reactions are
incorporated via Strang splitting~\cite{strang:1968}, giving a method
that is overall second-order accurate in space and time.
\castro\ uses the \amrex\ adaptive mesh refinement
library~\cite{amrex_joss} to manage a hierarchy of grids at different
resolutions.

@@

Since the neutron star rotates, we work in a corotating frame, with
angular velocity $\Omegab$.  Further, for the two-dimensional
simulations presented here, we work in axisymmetric coordinates,
but we advective a third component of velocity, coming out of the
simulation plane, that participates in the Coriolis force (sometimes
described as a 2.5D simulation).  We will take the
\castro\ $x$-coordinate to be the cylindrical radial coordinate with
corresponding velocity $u$, the \castro\ $y$-coordinate to be the
cylindrical vertical coordinate with corresponding velocity $v$, and the
\castro\ $z$-coordinate to be the cylindrical azimuthal coordinate,
with corresponding velocity $w$.  A righthanded coordinate system has
positive $w$ pointing out of the page.  We will take $\Omegab =
\Omega_0 \hat{\bf y}$ for the angular rotation rate, and $\gb = -g
\hat{\bf y}$ for the gravitational acceleration, with $g$ constant.
With these choices, the Coriolis force is:
\begin{equation}
-2\rho \Omegab \times \Ub =
   -2\rho \left ( \Omega_0 w \hat{x} - \Omega_0 u \hat{z} \right )
\end{equation}
We will neglect the centrifugal force---with our plane-parallel
geometry, this will act only in the lateral direction, and is not
expected to greatly affect the dynamics.  Carrying the Coriolis force
allows us to capture the geostrophic balance\MarginPar{ref} that sets
up via lateral hydrostatic equilibrium.

Writing the momentum equation in terms of the $u$, $v$, and $w$
components, and neglecting the centrifugal force, we have:
\begin{align}
\frac{\partial (\rho u)}{\partial t} + \nabla \cdot (\rho u \Ub) +
     \frac{\partial p}{\partial x} &= -2\rho \Omega_0 w  \\
%
\frac{\partial (\rho v)}{\partial t} + \nabla \cdot (\rho v \Ub) +
     \frac{\partial p}{\partial y} &= -\rho g \\
%
\frac{\partial (\rho w)}{\partial t} + \nabla \cdot (\rho w \Ub) +
     \cancelto{0}{\frac{\partial p}{\partial z}} &=
    2\rho \Omega_0 u 
\end{align}
where we cancel $\partial p/\partial z$ because there are no
variations in the azimuthal direction.  This allows us to recast the
$w$-velocity equation as a simple advection equation:
\begin{equation}
\frac{\partial w}{\partial t} + \Ub \cdot \nabla w = 2\Omega_0 u
\end{equation}
In our geometry, the flame will propagate from left to right, so $u$
will be positive and the Coriolis force results in $w > 0$ (out of the
simulation plane).  The algorithmic implementation of rotation in
\castro\ is described in \cite{wdmergerI}.

@@

\subsection{General Features}

\AssignTo{Mike, Kiran, Alice}

Here we'll show the large scale dynamics, colormap plots, etc. and discuss the trends.

\subsection{Flame Acceleration}

\AssignTo{Kiran}

Describe the diagnostics used to measure the flame speed and compare to laminar.

\AssignTo{Kiran, Mike}

\subsection{Entrainment, Flow Features}

\AssignTo{Alice, Don}

Show the phase plots and whatever it was that Don cooked up.

@@

\AssignTo{Mike, Kiran}

@@

\AssignTo{Mike, Kiran}

We look at iso7 instead of aprox13

\MarginPar{do we want to do a low res run?}

@@

\AssignTo{everyone}

@@

MHD, radiation transport, GPUs

@@

\software{AMReX \citep{amrex_joss},
          Castro \citep{castro},
          GCC (\url{https://gcc.gnu.org/}),
          linux (\url{https://www.kernel.org/}),
          matplotlib (\citealt{Hunter:2007}, \url{http://matplotlib.org/}),
          NumPy \citep{numpy,numpy2},
          python (\url{https://www.python.org/}),
          valgrind \citep{valgrind},
          VODE \citep{vode},
          yt \citep{yt}}
